Reanalysis Step 7: Alternative Propensity Score Specifications

### 2025-09-01 reanalysis-7-alternative_specifications.qmd
consistecy on ix

library(MatchIt)
library(WeightIt)
library(twang)
library(cobalt)
library(mice)
library(survey)
library(ggplot2)
library(gridExtra)

@@

cat(paste0("Running ", max_imputations_to_process, " of ", n_imputations, " imputations\n"))
```

@@

```{r define-methods}
# Define all PS methods to compare
ps_methods <- list(
  # MatchIt Methods - Nearest Neighbor
  list(name = "nearest_glm", package = "matchit", method = "nearest", 
       distance = "glm", estimand = "ATT", ratio = 1),

  list(name = "nearest_gam", package = "matchit", method = "nearest", 
       distance = "gam", estimand = "ATT", ratio = 1),

  list(name = "nearest_gbm", package = "matchit", method = "nearest", 
       distance = "gbm", estimand = "ATT", ratio = 1,
       distance.options = list(n.trees = 1000, interaction.depth = 3)),

  list(name = "nearest_lasso", package = "matchit", method = "nearest", 
       distance = "lasso", estimand = "ATT", ratio = 1),

  list(name = "nearest_rpart", package = "matchit", method = "nearest", 
       distance = "rpart", estimand = "ATT", ratio = 1),

  list(name = "mahalanobis", package = "matchit", method = "nearest", 
       distance = "mahalanobis", estimand = "ATT", ratio = 1),

@@

  list(name = "bart", package = "weightit", method = "bart", 
       estimand = "ATT"),

  # twang Method (already run, load results)
  list(name = "twang_gbm", package = "twang", method = "gbm", 
       estimand = "ATT", precomputed = TRUE)
)

# Create summary table of methods
methods_summary <- data.frame(
  Method = sapply(ps_methods, function(x) x$name),
  Package = sapply(ps_methods, function(x) x$package),
  Type = ifelse(grepl("nearest|mahalanobis", sapply(ps_methods, function(x) x$name)), 
                "Matching", 
                ifelse(grepl("subclass", sapply(ps_methods, function(x) x$name)),
                       "Subclassification", "IPTW")),
  stringsAsFactors = FALSE
)

kable(methods_summary, caption = "Propensity score methods to compare")
```

## Run Alternative Specifications

```{r run-alternatives}
#| warning: false
#| message: false

# Initialize results storage
all_results <- list()
balance_results <- list()
computation_times <- list()

# Load pre-computed twang results
twang_weights <- readRDS(file.path(reanalysis_data_dir, "twang_weights.rds"))

# Process each imputation
for (imp_num in 1:max_imputations_to_process) {
  cat(paste0("\n--- Processing Imputation ", imp_num, " ---\n"))

  imp_data <- imputed_datasets[[imp_num]]

  # Store results for this imputation
  imp_results <- list()

  # Process each method
  for (method_spec in ps_methods) {
    cat(paste0("  Method: ", method_spec$name, "... "))

    start_time <- Sys.time()

    tryCatch({
      if (method_spec$name == "twang_gbm") {
        # Use pre-computed twang weights
        weights <- twang_weights[[imp_num]]
        matched_data <- imp_data
        matched_data$weights <- weights
        balance_obj <- NULL  # Will calculate balance separately

      } else if (method_spec$package == "matchit") {
        # MatchIt methods
        match_args <- list(
          formula = ps_formula,
          data = imp_data,
          method = method_spec$method,
          estimand = method_spec$estimand
        )

        # Add distance specification if not Mahalanobis
        if (method_spec$name != "mahalanobis") {
          match_args$distance <- method_spec$distance
        }

        # Add method-specific options
        if (!is.null(method_spec$ratio)) {
          match_args$ratio <- method_spec$ratio
        }
        if (!is.null(method_spec$subclass)) {
          match_args$subclass <- method_spec$subclass
        }
        if (!is.null(method_spec$distance.options)) {
          match_args$distance.options <- method_spec$distance.options
        }

        # Run matching
        match_obj <- do.call(matchit, match_args)

        # Extract matched data and weights
        if (method_spec$method == "subclass") {
          matched_data <- match.data(match_obj, subclass = "subclass")
          matched_data$weights <- matched_data$weights * matched_data$subclass.weights
        } else {
          matched_data <- match.data(match_obj)
        }

        # Get balance statistics
        balance_obj <- bal.tab(match_obj, stats = c("m", "v"), thresholds = c(m = 0.1))

      } else if (method_spec$package == "weightit") {
        # WeightIt methods
        weight_obj <- weightit(
          formula = ps_formula,
          data = imp_data,
          method = method_spec$method,
          estimand = method_spec$estimand,
          over = ifelse(is.null(method_spec$over), FALSE, method_spec$over)
        )

        # Extract weights
        weights <- weight_obj$weights
        matched_data <- imp_data
        matched_data$weights <- weights

        # Get balance statistics
        balance_obj <- bal.tab(weight_obj, stats = c("m", "v"), thresholds = c(m = 0.1))
      }

      end_time <- Sys.time()
      comp_time <- as.numeric(difftime(end_time, start_time, units = "secs"))

      # Store results
      imp_results[[method_spec$name]] <- list(
        data = matched_data,
        weights = matched_data$weights,
        balance = balance_obj,
        computation_time = comp_time
      )

      cat(paste0("done (", round(comp_time, 2), "s)\n"))

    }, error = function(e) {
      cat(paste0("FAILED - ", e$message, "\n"))
      imp_results[[method_spec$name]] <- list(
        data = NULL,
        weights = NULL,
        balance = NULL,
        error = e$message,
        computation_time = NA
      )
    })
  }

  # Store imputation results
  all_results[[paste0("imp_", imp_num)]] <- imp_results
}

cat("\n✓ All methods processed\n")
```

## Fit Outcome Models

```{r fit-outcomes}
# Get the interaction formulas from functions.R
interaction_formulas <- get_analysis_formulas_reanalysis_interaction()

# Use the full formula which includes three-way interaction
outcome_formula <- interaction_formulas$full

# Store outcome model results
outcome_results <- list()

for (imp_num in 1:max_imputations_to_process) {
  imp_outcome_results <- list()

  for (method_name in names(all_results[[paste0("imp_", imp_num)]])) {
    method_result <- all_results[[paste0("imp_", imp_num)]][[method_name]]

    if (!is.null(method_result$data)) {
      # Fit weighted outcome model
      design <- svydesign(
        ids = ~ e_mrn_deidentified,
        weights = ~ weights,
        data = method_result$data
      )

      model <- svyglm(
        outcome_formula,
        design = design,
        family = quasibinomial()
      )

      # Extract lapse coefficient
      coef_summary <- coef(summary(model))
      lapse_coef <- coef_summary["ever_lapse_binary", ]

      imp_outcome_results[[method_name]] <- data.frame(
        method = method_name,
        imputation = imp_num,
        estimate = lapse_coef["Estimate"],
        se = lapse_coef["Std. Error"],
        or = exp(lapse_coef["Estimate"]),
        ci_lower = exp(lapse_coef["Estimate"] - 1.96 * lapse_coef["Std. Error"]),
        ci_upper = exp(lapse_coef["Estimate"] + 1.96 * lapse_coef["Std. Error"]),
        p_value = lapse_coef["Pr(>|t|)"]
      )
    }
  }

  outcome_results[[paste0("imp_", imp_num)]] <- do.call(rbind, imp_outcome_results)
}

# Combine all outcome results
all_outcome_results <- do.call(rbind, outcome_results)
row.names(all_outcome_results) <- NULL
```

## Pool Results Across Imputations

```{r pool-results}
# Pool results using Rubin's rules for each method
pooled_results <- list()

for (method_name in unique(all_outcome_results$method)) {
  method_data <- all_outcome_results[all_outcome_results$method == method_name, ]

  # Calculate pooled estimate (Rubin's rules)
  m <- nrow(method_data)  # Number of imputations

  # Pooled estimate
  q_bar <- mean(method_data$estimate)

  # Within-imputation variance
  u_bar <- mean(method_data$se^2)

  # Between-imputation variance
  b <- var(method_data$estimate)

  # Total variance
  t_var <- u_bar + (1 + 1/m) * b

  # Standard error
  se_pooled <- sqrt(t_var)

  # Degrees of freedom (Barnard-Rubin)
  lambda <- (1 + 1/m) * b / t_var
  df_old <- (m - 1) / lambda^2
  df_obs <- 90000  # Large sample approximation
  df_adj <- df_old * df_obs / (df_old + df_obs)

  # Confidence intervals
  t_crit <- qt(0.975, df_adj)
  ci_lower <- q_bar - t_crit * se_pooled
  ci_upper <- q_bar + t_crit * se_pooled

  # P-value
  t_stat <- q_bar / se_pooled
  p_value <- 2 * pt(-abs(t_stat), df_adj)

  pooled_results[[method_name]] <- data.frame(
    method = method_name,
    estimate = q_bar,
    se = se_pooled,
    or = exp(q_bar),
    ci_lower = exp(ci_lower),
    ci_upper = exp(ci_upper),
    p_value = p_value,
    fmi = lambda,
    n_imputations = m
  )
}

# Combine pooled results
pooled_df <- do.call(rbind, pooled_results)
row.names(pooled_df) <- NULL

# Sort by OR
pooled_df <- pooled_df[order(pooled_df$or), ]

# Add method type
pooled_df <- merge(pooled_df, methods_summary[, c("Method", "Type", "Package")], 
                   by.x = "method", by.y = "Method", all.x = TRUE)

# Display pooled results
kable(pooled_df[, c("method", "Type", "or", "ci_lower", "ci_upper", "p_value")],
      caption = "Pooled results across all methods",
      digits = 3,
      col.names = c("Method", "Type", "OR", "CI Lower", "CI Upper", "P-value"))
```

## Forest Plot Comparison

```{r forest-plot}
#| fig-height: 10
#| fig-width: 12

# Prepare data for forest plot
forest_data <- pooled_df

# Create method labels
forest_data$method_label <- paste0(
  forest_data$method, " (", forest_data$Package, ")"
)

# Order by OR
forest_data <- forest_data[order(forest_data$or), ]
forest_data$method_label <- factor(forest_data$method_label, 
                                  levels = forest_data$method_label)

# Define colors by package
forest_data$color <- case_when(
  forest_data$Package == "matchit" ~ "#2E86AB",
  forest_data$Package == "weightit" ~ "#A23B72",
  forest_data$Package == "twang" ~ "#F18F01",
  TRUE ~ "black"
)

# Create forest plot
forest_plot <- ggplot(forest_data, aes(x = method_label, y = or)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", size = 0.5) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper, color = color),
                  size = 0.8, linewidth = 0.8) +
  geom_point(aes(color = color), size = 3) +
  scale_color_identity() +
  coord_flip() +
  scale_y_continuous(
    breaks = seq(0.8, max(forest_data$ci_upper) + 0.2, by = 0.2),
    limits = c(0.8, max(forest_data$ci_upper) + 0.2)
  ) +
  labs(
    x = NULL,
    y = "Odds Ratio (95% CI)",
    title = "Comparison of Propensity Score Methods",
    subtitle = paste0("Effect of Lapsing on Vision Impairment/Blindness (", 
                     max_imputations_to_process, " imputations)"),
    caption = "Doubly robust estimation with ATT weights\nColors: Blue=MatchIt, Purple=WeightIt, Orange=twang"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 11),
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor = element_blank(),
    plot.margin = margin(20, 20, 20, 20),
    plot.caption = element_text(size = 9, hjust = 0.5, margin = margin(t = 15))
  )

# Add OR values as text
forest_plot <- forest_plot +
  geom_text(aes(label = sprintf("%.2f", or)), 
            hjust = -0.3, size = 3.5)

@@

# Save plot
ggsave(file.path(reanalysis_data_dir, "forest_plot_all_methods.png"),
       forest_plot, width = 12, height = 10, dpi = 300)
```

## Balance Assessment Across Methods

```{r balance-assessment}
#| fig-height: 8
#| fig-width: 12

# Extract balance statistics for first imputation (as example)
balance_stats <- list()

for (method_name in names(all_results$imp_1)) {
  method_result <- all_results$imp_1[[method_name]]

  if (!is.null(method_result$balance)) {
    bal_obj <- method_result$balance

    # Extract standardized mean differences
    smd_data <- bal_obj$Balance

    # Get max absolute SMD
    max_smd <- max(abs(smd_data$Diff.Adj), na.rm = TRUE)
    mean_smd <- mean(abs(smd_data$Diff.Adj), na.rm = TRUE)

    balance_stats[[method_name]] <- data.frame(
      method = method_name,
      max_smd = max_smd,
      mean_smd = mean_smd,
      n_balanced = sum(abs(smd_data$Diff.Adj) < 0.1, na.rm = TRUE),
      n_total = nrow(smd_data)
    )
  } else if (method_name == "twang_gbm") {
    # Calculate balance for twang manually
    imp_data <- imputed_datasets[[1]]
    imp_data$weights <- twang_weights[[1]]

    bal_obj <- bal.tab(
      ps_formula,
      data = imp_data,
      weights = imp_data$weights,
      estimand = "ATT",
      stats = c("m", "v"),
      thresholds = c(m = 0.1)
    )

    smd_data <- bal_obj$Balance
    max_smd <- max(abs(smd_data$Diff.Adj), na.rm = TRUE)
    mean_smd <- mean(abs(smd_data$Diff.Adj), na.rm = TRUE)

    balance_stats[[method_name]] <- data.frame(
      method = method_name,
      max_smd = max_smd,
      mean_smd = mean_smd,
      n_balanced = sum(abs(smd_data$Diff.Adj) < 0.1, na.rm = TRUE),
      n_total = nrow(smd_data)
    )
  }
}

# Combine balance statistics
balance_df <- do.call(rbind, balance_stats)
row.names(balance_df) <- NULL

# Add package info
balance_df <- merge(balance_df, methods_summary[, c("Method", "Package")],
                    by.x = "method", by.y = "Method", all.x = TRUE)

# Sort by mean SMD
balance_df <- balance_df[order(balance_df$mean_smd), ]

# Display balance table
kable(balance_df[, c("method", "max_smd", "mean_smd", "n_balanced", "n_total")],
      caption = "Balance statistics across methods (Imputation 1)",
      digits = 3,
      col.names = c("Method", "Max |SMD|", "Mean |SMD|", "N Balanced", "N Covariates"))

# Create balance comparison plot
balance_df$method <- factor(balance_df$method, levels = balance_df$method)

balance_plot <- ggplot(balance_df, aes(x = method)) +
  geom_col(aes(y = mean_smd, fill = Package), alpha = 0.7) +
  geom_point(aes(y = max_smd), size = 3, color = "red", shape = 17) +
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "red", alpha = 0.5) +
  coord_flip() +
  scale_fill_manual(values = c(matchit = "#2E86AB", weightit = "#A23B72", twang = "#F18F01")) +
  labs(
    x = NULL,
    y = "Standardized Mean Difference",
    title = "Covariate Balance Across Methods",
    subtitle = "Bars = Mean |SMD|, Red triangles = Max |SMD|, Dashed line = 0.1 threshold",
    fill = "Package"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    legend.position = "bottom"
  )

print(balance_plot)
```

## Love Plots for Selected Methods

```{r love-plots}
#| fig-height: 10
#| fig-width: 14

# Create Love plots for best performing methods from each package
selected_methods <- c("nearest_gbm", "entropy", "twang_gbm")

love_plots <- list()

for (method_name in selected_methods) {
  if (method_name %in% names(all_results$imp_1)) {

    # Get the data and weights
    method_data <- all_results$imp_1[[method_name]]$data

    if (!is.null(method_data)) {
      # Create Love plot
      love <- love.plot(
        bal.tab(ps_formula, 
                data = method_data,
                weights = method_data$weights,
                estimand = "ATT",
                s.d.denom = "treated"),
        thresholds = c(m = 0.1),
        title = paste("Love Plot:", method_name),
        colors = c("darkblue", "lightblue"),
        shapes = c(16, 17),
        line = TRUE
      ) +
        theme_minimal() +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"))

      love_plots[[method_name]] <- love
    }
  }
}

# Arrange Love plots
if (length(love_plots) > 0) {
  grid.arrange(grobs = love_plots, ncol = 2)
}
```

## Weight Distribution Comparison

```{r weight-distribution}
#| fig-height: 8
#| fig-width: 12

# Compare weight distributions for IPTW methods
iptw_methods <- c("cbps", "entropy", "bart", "twang_gbm")

weight_data <- list()

for (method_name in iptw_methods) {
  if (method_name %in% names(all_results$imp_1)) {
    weights <- all_results$imp_1[[method_name]]$weights

    if (!is.null(weights)) {
      weight_data[[method_name]] <- data.frame(
        method = method_name,
        weights = weights,
        treated = imputed_datasets[[1]]$ever_lapse_binary
      )
    }
  }
}

if (length(weight_data) > 0) {
  # Combine weight data
  weight_df <- do.call(rbind, weight_data)

  # Create weight distribution plot
  weight_dist_plot <- ggplot(weight_df, aes(x = weights, fill = factor(treated))) +
    geom_histogram(alpha = 0.6, position = "identity", bins = 50) +
    facet_wrap(~ method, scales = "free", ncol = 2) +
    scale_fill_manual(values = c("0" = "blue", "1" = "red"),
                     labels = c("0" = "Control", "1" = "Treated"),
                     name = "Group") +
    labs(
      x = "Weight",
      y = "Count",
      title = "Weight Distributions for IPTW Methods",
      subtitle = "Comparison across different weighting approaches"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5),
      legend.position = "bottom"
    )

  print(weight_dist_plot)

  # Weight statistics table
  weight_stats <- weight_df %>%
    group_by(method) %>%
    summarise(
      Mean = mean(weights),
      SD = sd(weights),
      Min = min(weights),
      Max = max(weights),
      `95th Pct` = quantile(weights, 0.95),
      `99th Pct` = quantile(weights, 0.99)
    )

  kable(weight_stats,
        caption = "Weight distribution statistics for IPTW methods",
        digits = 3)
}
```

## Computation Time Comparison

```{r computation-time}
# Extract computation times
comp_times <- list()

>>>>

library(MatchIt)
library(WeightIt)
library(twang)
library(cobalt)
library(mice)
library(survey)
library(ggplot2)
library(gridExtra)
library(parallel)
library(doParallel)
library(forestplot)

# Initialize logging
log_file <- init_log("alternative_ps_methods")

@@

# Caching configuration
use_cache <- TRUE
cache_dir <- file.path(reanalysis_data_dir, "ps_alternative_methods")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
  log_message(paste0("Created cache directory: ", cache_dir))
}

# Parallel processing configuration
max_cores_for_ps <- 2  # Limit cores for memory management

log_message(paste0("Analysis mode: ", analysis_mode))
log_message(paste0("Processing ", max_imputations_to_process, " of ", n_imputations, " imputations"))
log_message(paste0("Caching: ", ifelse(use_cache, "ENABLED", "DISABLED")))
```

@@

```{r define-methods}
# Define all PS methods to compare (excluding twang_gbm which we'll load from reanalysis-4)
ps_methods <- list(
  # MatchIt Methods - Nearest Neighbor with replacement
  list(name = "nearest_glm", package = "matchit", method = "nearest", 
       distance = "glm", estimand = "ATT", replace = TRUE),

  list(name = "nearest_gam", package = "matchit", method = "nearest", 
       distance = "gam", estimand = "ATT", replace = TRUE),

  list(name = "nearest_gbm", package = "matchit", method = "nearest", 
       distance = "gbm", estimand = "ATT", replace = TRUE,
       distance.options = list(n.trees = 1000, interaction.depth = 3)),

  list(name = "nearest_lasso", package = "matchit", method = "nearest", 
       distance = "lasso", estimand = "ATT", replace = TRUE),

  list(name = "nearest_rpart", package = "matchit", method = "nearest", 
       distance = "rpart", estimand = "ATT", replace = TRUE),

  list(name = "nearest_mahalanobis", package = "matchit", method = "nearest", 
       distance = "mahalanobis", estimand = "ATT", replace = TRUE),

@@

  list(name = "bart", package = "weightit", method = "bart", 
       estimand = "ATT")
)

# Note: twang_gbm results will be loaded from reanalysis-4

log_message(paste0("Defined ", length(ps_methods), " PS methods to compare"))
```

## Propensity Score Estimation Functions

```{r ps-functions}
# Function to process a single PS method for one imputation
process_ps_method <- function(imputation_id, data, method_params, ps_formula, cache_dir, use_cache = TRUE) {
  method_name <- method_params$name
  cache_file <- file.path(cache_dir, paste0(method_name, "_imp", imputation_id, ".rds"))

  # Check cache
  if (use_cache && file.exists(cache_file)) {
    cat(paste0("  Loading cached results for ", method_name, " (imp ", imputation_id, ")\n"))
    return(readRDS(cache_file))
  }

  cat(paste0("  Estimating ", method_name, " for imputation ", imputation_id, "\n"))

  result <- tryCatch({
    if (method_params$package == "matchit") {
      # MatchIt methods
      args <- list(
        formula = ps_formula,
        data = data,
        method = method_params$method,
        estimand = method_params$estimand
      )

      # Add method-specific parameters
      if (!is.null(method_params$distance)) args$distance <- method_params$distance
      if (!is.null(method_params$replace)) args$replace <- method_params$replace
      if (!is.null(method_params$subclass)) args$subclass <- method_params$subclass
      if (!is.null(method_params$distance.options)) args$distance.options <- method_params$distance.options

      do.call(matchit, args)

    } else if (method_params$package == "weightit") {
      # WeightIt methods
      args <- list(
        formula = ps_formula,
        data = data,
        method = method_params$method,
        estimand = method_params$estimand
      )

      # Add method-specific parameters
      if (!is.null(method_params$over)) args$over <- method_params$over

      do.call(weightit, args)

    } else {
      stop(paste("Unknown package:", method_params$package))
    }
  }, error = function(e) {
    cat(paste0("  ERROR in ", method_name, " (imp ", imputation_id, "): ", e$message, "\n"))
    return(NULL)
  })

  # Save to cache
  if (!is.null(result) && use_cache) {
    saveRDS(result, cache_file)
  }

  return(result)
}

# Function to process all imputations for a single method
process_method_all_imputations <- function(method_params, imputed_datasets, ps_formula, 
                                          cache_dir, use_cache = TRUE, n_cores = 1) {
  n_imp <- length(imputed_datasets)
  method_name <- method_params$name

  log_message(paste0("Processing method: ", method_name, " across ", n_imp, " imputations"))

  if (n_cores > 1 && .Platform$OS.type == "unix") {
    # Parallel processing for Unix/Mac
    results <- mclapply(
      1:n_imp,
      function(i) process_ps_method(i, imputed_datasets[[i]], method_params, ps_formula, cache_dir, use_cache),
      mc.cores = n_cores
    )
  } else if (n_cores > 1 && .Platform$OS.type != "unix") {
    # Parallel processing for Windows
    cl <- makeCluster(n_cores)
    clusterExport(cl, c("process_ps_method", "ps_formula", "cache_dir", "use_cache", "method_params"),
                  envir = environment())
    clusterEvalQ(cl, {
      library(MatchIt)
      library(WeightIt)
    })

    results <- parLapply(cl, 1:n_imp, function(i) {
      process_ps_method(i, imputed_datasets[[i]], method_params, ps_formula, cache_dir, use_cache)
    })

    stopCluster(cl)
  } else {
    # Sequential processing
    results <- lapply(
      1:n_imp,
      function(i) process_ps_method(i, imputed_datasets[[i]], method_params, ps_formula, cache_dir, use_cache)
    )
  }

  return(results)
}
```

## Run Propensity Score Methods

```{r run-ps-methods}
# Process only the specified number of imputations
imputed_datasets_subset <- imputed_datasets[1:min(max_imputations_to_process, length(imputed_datasets))]

# Initialize results storage
all_ps_results <- list()

# Process each method sequentially (to manage memory)
for (method_params in ps_methods) {
  method_name <- method_params$name
  log_message(paste0("\n", paste(rep("=", 50), collapse = "")))
  log_message(paste0("Starting method: ", method_name))

  # Process all imputations for this method
  method_results <- process_method_all_imputations(
    method_params = method_params,
    imputed_datasets = imputed_datasets_subset,
    ps_formula = ps_formula,
    cache_dir = cache_dir,
    use_cache = use_cache,
    n_cores = max_cores_for_ps
  )

  all_ps_results[[method_name]] <- method_results

  # Clean up memory after each method
  gc()

  log_message(paste0("Completed method: ", method_name))
}

log_message("All PS methods completed")
```

## Load twang_gbm Results from Reanalysis-4

```{r load-twang}
# Load pre-computed twang_gbm results
twang_cache_dir <- file.path(reanalysis_data_dir, "ps_cache")
twang_results <- list()

for (i in 1:max_imputations_to_process) {
  twang_file <- file.path(twang_cache_dir, paste0("ps_imp_", i, ".rds"))
  if (file.exists(twang_file)) {
    twang_results[[i]] <- readRDS(twang_file)
    log_message(paste0("Loaded twang_gbm results for imputation ", i))
  } else {
    log_message(paste0("WARNING: twang_gbm results not found for imputation ", i))
    twang_results[[i]] <- NULL
  }
}

# Add to all results
all_ps_results[["twang_gbm"]] <- twang_results

log_message(paste0("Loaded twang_gbm results for ", sum(!sapply(twang_results, is.null)), " imputations"))
```

## Outcome Analysis

```{r outcome-analysis}
# Get outcome formula from functions.R
outcome_formulas <- get_analysis_formulas_reanalysis()
outcome_formula <- outcome_formulas$full

log_message("Starting outcome analysis for all methods")

# Function to fit outcome model for a single method and imputation
fit_outcome_model <- function(ps_result, data, formula, method_name) {
  if (is.null(ps_result)) {
    return(NULL)
  }

  tryCatch({
    if (method_name == "twang_gbm") {
      # twang weights
      weights <- ps_result$w[, "es.mean.ATT"]
      data$weights <- weights
      design <- svydesign(ids = ~e_mrn_deidentified, weights = ~weights, data = data)

    } else if (inherits(ps_result, "matchit")) {
      # MatchIt results
      if (grepl("subclass", method_name)) {
        # Subclassification
        matched_data <- match.data(ps_result, subclass = "subclass")
        design <- svydesign(ids = ~e_mrn_deidentified, data = matched_data, weights = ~weights)
        design <- postStratify(design, ~subclass, xtabs(~subclass, data = matched_data))
      } else {
        # Matching
        matched_data <- get_matches(ps_result)
        design <- svydesign(ids = ~e_mrn_deidentified, data = matched_data, weights = ~1)
      }

    } else if (inherits(ps_result, "weightit")) {
      # WeightIt results
      weights <- ps_result$weights
      data$weights <- weights
      design <- svydesign(ids = ~e_mrn_deidentified, weights = ~weights, data = data)

    } else {
      stop("Unknown result type")
    }

    # Fit the model
    model <- svyglm(formula, design = design, family = quasibinomial())

    # Extract coefficient for lapse
    coef_lapse <- coef(summary(model))["ever_lapse_binary", ]

    return(list(
      estimate = coef_lapse["Estimate"],
      se = coef_lapse["Std. Error"],
      model = model
    ))

  }, error = function(e) {
    cat(paste0("  Error fitting model for ", method_name, ": ", e$message, "\n"))
    return(NULL)
  })
}

# Process outcome models for all methods and imputations
outcome_results <- list()

for (method_name in names(all_ps_results)) {
  log_message(paste0("Fitting outcome models for ", method_name))

  method_outcomes <- list()
  ps_results <- all_ps_results[[method_name]]

  for (i in 1:length(ps_results)) {
    if (!is.null(ps_results[[i]])) {
      outcome <- fit_outcome_model(
        ps_result = ps_results[[i]],
        data = imputed_datasets_subset[[i]],
        formula = outcome_formula,
        method_name = method_name
      )
      method_outcomes[[i]] <- outcome
    }
  }

  outcome_results[[method_name]] <- method_outcomes
}

log_message("Outcome analysis completed")
```

## Pool Results Using Rubin's Rules

```{r pool-results}
# Function to pool results across imputations using Rubin's rules
pool_results <- function(method_outcomes) {
  # Filter out NULL results
  valid_outcomes <- method_outcomes[!sapply(method_outcomes, is.null)]

  if (length(valid_outcomes) == 0) {
    return(NULL)
  }

  # Extract estimates and SEs
  estimates <- sapply(valid_outcomes, function(x) x$estimate)
  ses <- sapply(valid_outcomes, function(x) x$se)

  # Remove any NA values
  valid_idx <- !is.na(estimates) & !is.na(ses)
  estimates <- estimates[valid_idx]
  ses <- ses[valid_idx]

  if (length(estimates) == 0) {
    return(NULL)
  }

  m <- length(estimates)  # Number of imputations

  # Pooled estimate (average of estimates)
  pooled_estimate <- mean(estimates)

  # Within-imputation variance (average of squared SEs)
  within_var <- mean(ses^2)

  # Between-imputation variance
  between_var <- var(estimates)

  # Total variance using Rubin's rules
  total_var <- within_var + between_var + (between_var / m)
  pooled_se <- sqrt(total_var)

  # Degrees of freedom (Barnard-Rubin adjustment)
  lambda <- (between_var + between_var/m) / total_var
  df_old <- (m - 1) / lambda^2

  # For large samples, use df_old
  df <- df_old

  # Calculate confidence intervals and p-value
  t_stat <- pooled_estimate / pooled_se
  p_value <- 2 * pt(abs(t_stat), df = df, lower.tail = FALSE)
  ci_lower <- pooled_estimate - qt(0.975, df) * pooled_se
  ci_upper <- pooled_estimate + qt(0.975, df) * pooled_se

  # Convert to odds ratio scale
  or <- exp(pooled_estimate)
  or_ci_lower <- exp(ci_lower)
  or_ci_upper <- exp(ci_upper)

  return(list(
    estimate = pooled_estimate,
    se = pooled_se,
    or = or,
    or_ci_lower = or_ci_lower,
    or_ci_upper = or_ci_upper,
    p_value = p_value,
    n_imputations = m,
    df = df
  ))
}

# Pool results for each method
pooled_results <- list()

for (method_name in names(outcome_results)) {
  pooled <- pool_results(outcome_results[[method_name]])
  if (!is.null(pooled)) {
    pooled$method <- method_name
    pooled_results[[method_name]] <- pooled
    log_message(paste0("Pooled results for ", method_name, ": OR = ", 
                      round(pooled$or, 3), " (", round(pooled$or_ci_lower, 3), 
                      ", ", round(pooled$or_ci_upper, 3), ")"))
  } else {
    log_message(paste0("WARNING: Could not pool results for ", method_name))
  }
}

# Save pooled results
saveRDS(pooled_results, file.path(reanalysis_data_dir, "alternative_ps_results_pooled.rds"))
log_message("Saved pooled results")
```

## Create Forest Plot

```{r forest-plot}
# Prepare data for forest plot
forest_data <- do.call(rbind, lapply(names(pooled_results), function(method) {
  res <- pooled_results[[method]]
  data.frame(
    method = method,
    or = res$or,
    lower = res$or_ci_lower,
    upper = res$or_ci_upper,
    p_value = res$p_value,
    stringsAsFactors = FALSE
  )
}))

# Order by odds ratio
forest_data <- forest_data[order(forest_data$or), ]

# Create more readable method names
forest_data$method_label <- gsub("_", " ", forest_data$method)
forest_data$method_label <- gsub("nearest ", "NN-", forest_data$method_label)
forest_data$method_label <- gsub("subclass ", "Subclass-", forest_data$method_label)
forest_data$method_label <- tools::toTitleCase(forest_data$method_label)

# Create forest plot using ggplot2
library(ggplot2)

forest_plot <- ggplot(forest_data, aes(x = or, y = reorder(method_label, or))) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.2) +
  geom_point(size = 3, shape = 18) +
  scale_x_continuous(
    trans = "log",
    breaks = c(0.5, 0.75, 1, 1.5, 2, 3),
    labels = c("0.5", "0.75", "1.0", "1.5", "2.0", "3.0")
  ) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "Propensity Score Method",
    title = "Effect of Lapse on Vision Loss Across PS Methods",
    subtitle = paste0("Based on ", max_imputations_to_process, " imputations with Rubin's pooling")
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.y = element_text(hjust = 0)
  )

@@

# Save plot
ggsave(
  filename = file.path(reanalysis_data_dir, "forest_plot_ps_methods.png"),
  plot = forest_plot,
  width = 10,
  height = 8,
  dpi = 300
)

log_message("Forest plot saved")
```

## Summary Tables

```{r summary-tables}
# Create detailed summary table
summary_table <- do.call(rbind, lapply(names(pooled_results), function(method) {
  res <- pooled_results[[method]]
  data.frame(
    Method = method,
    `OR` = sprintf("%.3f", res$or),
    `95% CI` = sprintf("(%.3f, %.3f)", res$or_ci_lower, res$or_ci_upper),
    `P-value` = sprintf("%.4f", res$p_value),
    `N Imputations` = res$n_imputations,
    stringsAsFactors = FALSE
  )
}))

# Sort by OR
summary_table <- summary_table[order(as.numeric(gsub("([0-9.]+)", "\\1", summary_table$OR))), ]

kable(summary_table, 
      caption = "Pooled Effect Estimates Across Propensity Score Methods",
      col.names = c("Method", "Odds Ratio", "95% CI", "P-value", "N Imputations"),
      align = c("l", "r", "r", "r", "r"))

# Save to CSV
write.csv(summary_table, 
          file.path(reanalysis_data_dir, "ps_methods_comparison_table.csv"),
          row.names = FALSE)

log_message("Summary table saved")

# Create comparison with twang_gbm as reference
if ("twang_gbm" %in% names(pooled_results)) {
  ref_or <- pooled_results[["twang_gbm"]]$or

  comparison_table <- do.call(rbind, lapply(names(pooled_results), function(method) {
    res <- pooled_results[[method]]
    data.frame(
      Method = method,
      OR = res$or,
      `Relative to twang` = sprintf("%.1f%%", 100 * (res$or / ref_or - 1)),
      stringsAsFactors = FALSE
    )
  }))

  kable(comparison_table,
        caption = "Comparison of Methods Relative to twang_gbm",
        col.names = c("Method", "Odds Ratio", "% Difference from twang_gbm"),
        align = c("l", "r", "r"))
}
```

## Summary

This analysis compared `r length(pooled_results)` different propensity score methods across `r max_imputations_to_process` multiply imputed datasets. Results were pooled using Rubin's rules to account for between-imputation variability.

Key findings:
- Method with smallest effect: `r names(pooled_results)[which.min(sapply(pooled_results, function(x) x$or))]`
- Method with largest effect: `r names(pooled_results)[which.max(sapply(pooled_results, function(x) x$or))]`
- Range of odds ratios: `r sprintf("%.3f to %.3f", min(sapply(pooled_results, function(x) x$or)), max(sapply(pooled_results, function(x) x$or)))`
